---
title: "2 comparing climate data"
author: "Matthew Ross"
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(RcppRoll)

gldas_dir = 'data/GLDAS/'

if(!file.exists(gldas_dir)){
  dir.create('data')
  dir.create(gldas_dir)
}
```


# Assignment


For this assignment we are going to compare climate data from a single point 
versus a watershed averaged climate data. We will be working over the Yampa river
watershed, one of the last undammed watersheds in the USA. 

## Point-climate acquisition. 

Using the 1_climate_data_yojoa.Rmd as an example download at least two types of climate (wind/temp/etc...) GLDAS data for the
Yampa River above Elkhead Creek. This is the site of a USGS gage that has data
dating back to 2004. 

```{r}
site_info <- tibble(site_no = '09244490',
                    lat = 40.5180278,
                    long = -107.3997838,
                    name = 'Yampa_hayden')
```

```{r}
#set parameters of function - these are found on the data rods website linked above
gldas_mod = c('GLDAS_NOAH025_3H_v2.0', 'GLDAS_NOAH025_3H_v2.1')
params = c('Wind_f_inst', 'Tair_f_inst')

#time period of interest - v 2.0 and v 2.1 have different time periods of coverage, and to keep the download from hanging, you have to have separate start/end dates
start_date_2.1 = '2004-01-01'
end_date_2.1 = '2023-01-01'

#GLDAS data are in 0.25 degree increments, of which Yojoa intersects with a few kernels, the following degrees would prioritize the kernel over the most water area of Yojoa
lat = 40.5180278
lon = -107.3997838

#function to make wwws to ping
make_www_2.1 = function(model, var){#, s_d, e_d, lat, lon) {
  s_d = start_date_2.1
  e_d = end_date_2.1
  paste0('https://hydro1.gesdisc.eosdis.nasa.gov/daac-bin/access/timeseries.cgi?variable=GLDAS2:', model, ':', var, '&startDate=', s_d, 'T00:00&endDate=', e_d, 'T21:00&location=GEOM:POINT(', lon, ',%20', lat, ')&type=asc2')
}

#make a list of all wwws for download
v2.1_www = unlist(map2(rep(gldas_mod[2], times = length(params)), params, make_www_2.1))
```

```{r, eval = F}
for(w21 in 1:length(v2.1_www)) {
  download.file(url = v2.1_www[w21], destfile = file.path(gldas_dir, paste0(rep(gldas_mod[2], length(params))[w21], '_', params[w21], '.csv')))
}
```

## Watershed averaged climate data

Using climate engine, download the two same parameters but for the watershed.
The watershed is stored as a .geojson in the yampa folder. Note, you likely need
to convert the watershed to a ".shp" file before getting the data from climate 
engine. 
```{r}
# Convert .geojson to .shp
library(sf)
library(mapview)

# Read the GeoJSON file
yampa <- st_read("data/yampa/yampa.geojson")
mapview(yampa)

# Write the data to a Shapefile
st_write(yampa[2,], "data/yampa/yampa.shp")
```


## Compare your climate data anyway you want

Make at least two plots comparing your point versus watershed-averaged climate data. 




